---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

1. Downloading file from Polish National Library database. Editing in MarcEdit: tools->marc processing tools -> marc split. Dividing for 200000 records and process.
2. Marc tools -> MarcBreaker and change from .mrc to .mrk; set default character encoding to UTF-8 (30 seconds per a text file that can be uploaded to R)
3. For loop in R (need to divide for a smaller parts)

```{r BN marc processing to table}
options(java.parameters = "-Xmx32000m")
pacman::p_load(plyr,stringr,tidyverse,reshape2,zoo,sqldf,splitstackshape,dplyr,svMisc,googlesheets)

#1
data_full <- data.frame()
x <- 1:10

#download database in pieces
for (i in x) {
progress(match(i,x), max.value = length(x))
b <- sprintf("%02d", i-1)
path1 <- "C:/Users/User/Desktop/bn_all_1.10.2019/msplit000000"
path2 <- b
path3 <- ".mrk"
path <- paste(path1,path2,path3,sep = "")
bn1 <- readLines(path, encoding = "UTF-8")
data1 <- data.frame(bn1)
remove(bn1)
data1 <- data1 %>%
  mutate(czy = grepl("^\\=LDR",bn1),
         row = ifelse(grepl("^\\=LDR",bn1),seq.int(nrow(data1)),NA)) %>%
  filter(bn1!="") %>%
  fill(row) %>%
  mutate(id = ifelse(grepl("^\\=009",bn1),str_replace_all(bn1,"(^\\=009 .*?\\w)(.*?)($)","\\2"),NA)) %>%
  group_by(row) %>%
  fill(id) %>%
  fill(id,.direction = "up") %>%
  ungroup() %>%
  select(-2,-3) %>%
  mutate(field = str_replace_all(bn1, "(=)(\\w{3})(.*)", "\\2"),
         content = str_replace_all(bn1,"(=)(\\w{3})(\\s{2})(.*)","\\4")) %>%
  select(2,3,4) %>%
  mutate(id_field = paste(id,field,sep = "|"))

count <- data1 %>%
  select(4,3)
count <- as.data.frame(table(count$id_field))
data1 <- merge(x=data1,y=count, by.x = "id_field", by.y = "Var1")
remove(count)
data1_to_join <- data1 %>%
  filter(Freq > 1)
data1_to_join <- ddply(data1_to_join, .(id_field), summarize, content = paste(content, collapse="|"))
data1_to_join <- mutate(data1_to_join,
                        id = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\1"),
                        field = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\3"))
data1_ok <- data1 %>%
  filter(Freq == 1) %>%
  select(-5) %>%
  select(1,4,2,3)
data1 <- rbind(data1_ok,data1_to_join)
remove(data1_ok)
remove(data1_to_join)
data1 <- data1[order(as.character(data1$id_field)),]
data1 <- data1 %>%
  select(3,4,2)
data_full <- rbind(data_full,data1)

}

write.csv2(data_full, file = "C:/Users/User/Desktop/bn_all_1.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

#2
data_full <- data.frame()

x <- 11:20

#download database in pieces
for (i in x) {
progress(match(i,x), max.value = length(x))
b <- sprintf("%02d", i-1)
path1 <- "C:/Users/User/Desktop/bn_all_1.10.2019/msplit000000"
path2 <- b
path3 <- ".mrk"
path <- paste(path1,path2,path3,sep = "")
bn1 <- readLines(path, encoding = "UTF-8")
data1 <- data.frame(bn1)
remove(bn1)
data1 <- data1 %>%
  mutate(czy = grepl("^\\=LDR",bn1),
         row = ifelse(grepl("^\\=LDR",bn1),seq.int(nrow(data1)),NA)) %>%
  filter(bn1!="") %>%
  fill(row) %>%
  mutate(id = ifelse(grepl("^\\=009",bn1),str_replace_all(bn1,"(^\\=009 .*?\\w)(.*?)($)","\\2"),NA)) %>%
  group_by(row) %>%
  fill(id) %>%
  fill(id,.direction = "up") %>%
  ungroup() %>%
  select(-2,-3) %>%
  mutate(field = str_replace_all(bn1, "(=)(\\w{3})(.*)", "\\2"),
         content = str_replace_all(bn1,"(=)(\\w{3})(\\s{2})(.*)","\\4")) %>%
  select(2,3,4) %>%
  mutate(id_field = paste(id,field,sep = "|"))

count <- data1 %>%
  select(4,3)
count <- as.data.frame(table(count$id_field))
data1 <- merge(x=data1,y=count, by.x = "id_field", by.y = "Var1")
remove(count)
data1_to_join <- data1 %>%
  filter(Freq > 1)
data1_to_join <- ddply(data1_to_join, .(id_field), summarize, content = paste(content, collapse="|"))
data1_to_join <- mutate(data1_to_join,
                        id = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\1"),
                        field = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\3"))
data1_ok <- data1 %>%
  filter(Freq == 1) %>%
  select(-5) %>%
  select(1,4,2,3)
data1 <- rbind(data1_ok,data1_to_join)
remove(data1_ok)
remove(data1_to_join)
data1 <- data1[order(as.character(data1$id_field)),]
data1 <- data1 %>%
  select(3,4,2)
data_full <- rbind(data_full,data1)

}

write.csv2(data_full, file = "C:/Users/User/Desktop/bn_all_2.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

#3
data_full <- data.frame()

x <- 21:30

#download database in pieces
for (i in x) {
progress(match(i,x), max.value = length(x))
b <- sprintf("%02d", i-1)
path1 <- "C:/Users/User/Desktop/bn_all_1.10.2019/msplit000000"
path2 <- b
path3 <- ".mrk"
path <- paste(path1,path2,path3,sep = "")
bn1 <- readLines(path, encoding = "UTF-8")
data1 <- data.frame(bn1)
remove(bn1)
data1 <- data1 %>%
  mutate(czy = grepl("^\\=LDR",bn1),
         row = ifelse(grepl("^\\=LDR",bn1),seq.int(nrow(data1)),NA)) %>%
  filter(bn1!="") %>%
  fill(row) %>%
  mutate(id = ifelse(grepl("^\\=009",bn1),str_replace_all(bn1,"(^\\=009 .*?\\w)(.*?)($)","\\2"),NA)) %>%
  group_by(row) %>%
  fill(id) %>%
  fill(id,.direction = "up") %>%
  ungroup() %>%
  select(-2,-3) %>%
  mutate(field = str_replace_all(bn1, "(=)(\\w{3})(.*)", "\\2"),
         content = str_replace_all(bn1,"(=)(\\w{3})(\\s{2})(.*)","\\4")) %>%
  select(2,3,4) %>%
  mutate(id_field = paste(id,field,sep = "|"))

count <- data1 %>%
  select(4,3)
count <- as.data.frame(table(count$id_field))
data1 <- merge(x=data1,y=count, by.x = "id_field", by.y = "Var1")
remove(count)
data1_to_join <- data1 %>%
  filter(Freq > 1)
data1_to_join <- ddply(data1_to_join, .(id_field), summarize, content = paste(content, collapse="|"))
data1_to_join <- mutate(data1_to_join,
                        id = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\1"),
                        field = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\3"))
data1_ok <- data1 %>%
  filter(Freq == 1) %>%
  select(-5) %>%
  select(1,4,2,3)
data1 <- rbind(data1_ok,data1_to_join)
remove(data1_ok)
remove(data1_to_join)
data1 <- data1[order(as.character(data1$id_field)),]
data1 <- data1 %>%
  select(3,4,2)
data_full <- rbind(data_full,data1)

}

write.csv2(data_full, file = "C:/Users/User/Desktop/bn_all_3.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

#4
data_full <- data.frame()

x <- 31:33
#download database in pieces
for (i in x) {
progress(match(i,x), max.value = length(x))
b <- sprintf("%02d", i-1)
path1 <- "C:/Users/User/Desktop/bn_all_1.10.2019/msplit000000"
path2 <- b
path3 <- ".mrk"
path <- paste(path1,path2,path3,sep = "")
bn1 <- readLines(path, encoding = "UTF-8")
data1 <- data.frame(bn1)
remove(bn1)
data1 <- data1 %>%
  mutate(czy = grepl("^\\=LDR",bn1),
         row = ifelse(grepl("^\\=LDR",bn1),seq.int(nrow(data1)),NA)) %>%
  filter(bn1!="") %>%
  fill(row) %>%
  mutate(id = ifelse(grepl("^\\=009",bn1),str_replace_all(bn1,"(^\\=009 .*?\\w)(.*?)($)","\\2"),NA)) %>%
  group_by(row) %>%
  fill(id) %>%
  fill(id,.direction = "up") %>%
  ungroup() %>%
  select(-2,-3) %>%
  mutate(field = str_replace_all(bn1, "(=)(\\w{3})(.*)", "\\2"),
         content = str_replace_all(bn1,"(=)(\\w{3})(\\s{2})(.*)","\\4")) %>%
  select(2,3,4) %>%
  mutate(id_field = paste(id,field,sep = "|"))

count <- data1 %>%
  select(4,3)
count <- as.data.frame(table(count$id_field))
data1 <- merge(x=data1,y=count, by.x = "id_field", by.y = "Var1")
remove(count)
data1_to_join <- data1 %>%
  filter(Freq > 1)
data1_to_join <- ddply(data1_to_join, .(id_field), summarize, content = paste(content, collapse="|"))
data1_to_join <- mutate(data1_to_join,
                        id = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\1"),
                        field = str_replace_all(data1_to_join$id_field,"(.*)(\\|)(.*)", "\\3"))
data1_ok <- data1 %>%
  filter(Freq == 1) %>%
  select(-5) %>%
  select(1,4,2,3)
data1 <- rbind(data1_ok,data1_to_join)
remove(data1_ok)
remove(data1_to_join)
data1 <- data1[order(as.character(data1$id_field)),]
data1 <- data1 %>%
  select(3,4,2)
data_full <- rbind(data_full,data1)

}

write.csv2(data_full, file = "C:/Users/User/Desktop/bn_all_4.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')


```

```{r }




bn1 <- read.csv2("C:/Users/User/Desktop/bn_data_all_1.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE) 
bn2 <- read.csv2("C:/Users/User/Desktop/bn_data_all_2.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE) 
bn3 <- read.csv2("C:/Users/User/Desktop/bn_data_all_3.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE) 

bn1 <- as.data.frame(bn1)
bn2 <- as.data.frame(bn2)
bn3 <- as.data.frame(bn3)

data_full <- rbind(bn1,bn2,bn3)

write.csv2(data_full, file = "C:/Users/User/Desktop/bn_data_all.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
data_full <- read.csv2("C:/Users/User/Desktop/bn_kartoteka_wzorcowa.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE) 

f.agg <- function(x) paste(x, collapse = "~")
data_full2 <- dcast(data_full, id ~ field, value.var="content", fun.aggregate = f.agg)

data_full2 <- dcast(data_full, id ~ field, value.var="content")
str_replace(data_full2$id,"(^.*?)(\\~.*?$)","\\1")
reshape2::dcast()
data_full2
write.csv2(data_full, file = "C:/Users/User/Desktop/bn_data_all_table.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

```

```{r ogarnianie kartoteki wzorcowe}
data_full2 <- data_full2 %>%
  select(-`665<U+2028>`)

data_full2$id <- str_replace(data_full2$id,"(^.*?)(\\~.*?$)","\\1")
data_full2$`001` <- str_replace(data_full2$`001`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`005` <- str_replace(data_full2$`005`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`008` <- str_replace(data_full2$`008`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`009` <- str_replace(data_full2$`009`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`010` <- str_replace(data_full2$`010`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`020` <- str_replace(data_full2$`020`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`022` <- str_replace(data_full2$`022`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`024` <- str_replace(data_full2$`024`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`034` <- str_replace(data_full2$`034`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`035` <- str_replace(data_full2$`035`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`040` <- str_replace(data_full2$`040`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`043` <- str_replace(data_full2$`043`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`045` <- str_replace(data_full2$`045`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`046` <- str_replace(data_full2$`046`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`072` <- str_replace(data_full2$`072`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`080` <- str_replace(data_full2$`080`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`100` <- str_replace(data_full2$`100`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`110` <- str_replace(data_full2$`110`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`111` <- str_replace(data_full2$`111`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`130` <- str_replace(data_full2$`130`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`150` <- str_replace(data_full2$`150`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`151` <- str_replace(data_full2$`151`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`155` <- str_replace(data_full2$`155`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`368` <- str_replace(data_full2$`368`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`370` <- str_replace(data_full2$`370`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`371` <- str_replace(data_full2$`371`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`372` <- str_replace(data_full2$`372`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`373` <- str_replace(data_full2$`373`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`374` <- str_replace(data_full2$`374`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`375` <- str_replace(data_full2$`375`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`376` <- str_replace(data_full2$`376`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`377` <- str_replace(data_full2$`377`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`378` <- str_replace(data_full2$`378`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`382` <- str_replace(data_full2$`382`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`388` <- str_replace(data_full2$`388`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`400` <- str_replace(data_full2$`400`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`410` <- str_replace(data_full2$`410`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`411` <- str_replace(data_full2$`411`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`430` <- str_replace(data_full2$`430`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`450` <- str_replace(data_full2$`450`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`451` <- str_replace(data_full2$`451`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`455` <- str_replace(data_full2$`455`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`500` <- str_replace(data_full2$`500`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`510` <- str_replace(data_full2$`510`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`511` <- str_replace(data_full2$`511`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`530` <- str_replace(data_full2$`530`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`550` <- str_replace(data_full2$`550`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`551` <- str_replace(data_full2$`551`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`555` <- str_replace(data_full2$`555`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`640` <- str_replace(data_full2$`640`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`641` <- str_replace(data_full2$`641`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`642` <- str_replace(data_full2$`642`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`643` <- str_replace(data_full2$`643`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`665` <- str_replace(data_full2$`665`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`666` <- str_replace(data_full2$`666`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`670` <- str_replace(data_full2$`670`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`675` <- str_replace(data_full2$`675`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`680` <- str_replace(data_full2$`680`,"(^.*?)(\\~.*?$)","\\1")
data_full2$`682` <- str_replace(data_full2$`682`,"(^.*?)(\\~.*?$)","\\1")
data_full2$LDR <- str_replace(data_full2$LDR,"(^.*?)(\\~.*?$)","\\1")
write.csv2(data_full2, file = "C:/Users/User/Desktop/bn_kartoteka_wzorcowa.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

```

2. Przygotowanie list książek dla danego roku

```{r wydobycie list ksiazek}
data_full <- read.csv2("C:/Users/User/Desktop/bn_data_all_table.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE)
data_full <- mutate(data_full,
                rok = str_sub(data_full$X008,8,11))
data_full <- data_full %>%
  filter(rok>2003&rok<2009)
write.csv2(data_full, file = "C:/Users/User/Desktop/bn_data_2004-2008.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
bn_ks <- mutate(data_full,
                czy_ks = (grepl("Ksi??ki", data_full$X380,ignore.case = TRUE)|data_full$X380=="")&substr(data_full$LDR,7,8)=="am")
bn_ks <- bn_ks %>%
  filter(czy_ks==TRUE) %>%
  select(-length(bn_ks)) %>%
  unique()
test <- sqldf("select X245, X260,X380,X300,LDR
              from bn_ks a")
write.csv2(bn_ks, file = "C:/Users/User/Desktop/bn_ks_2004-2008.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
bn_ks <- read.csv2("C:/Users/User/Desktop/bn_ks_2004-2008.csv", encoding = "Windows-1250", header = TRUE, stringsAsFactors = FALSE) 


bn_ks_2005_2008 <- bn_ks %>%
  filter(rok>2004&rok<2009)

#UKD_lit
UKD_lit <- mutate(bn_ks_2005_2008,
                  ukd_lit = grepl("\\$a8", bn_ks_2005_2008$X080))
UKD_lit <- UKD_lit %>%
  filter(ukd_lit == TRUE)

#gat_lit_bez_UKD_lit
gatunki_litX655 <- data.frame(X655 = UKD_lit$X655) %>%
  unique() %>%
  filter(!is.na(X655))
gatunki_litX655$X655 <-str_replace_all(gatunki_litX655$X655,"(^|\\|)","~\\1")
gatunki_litX655 <- cSplit(gatunki_litX655,"X655",sep = "~",direction = "long")
gatunki_litX655 <- gatunki_litX655%>%
  filter(X655!="")
gatunki_litX655$X655 <- str_replace_all(gatunki_litX655$X655,"(.*?)(\\$a.*)(\\$2.*)","\\2")
gatunki_litX655 <- gatunki_litX655 %>%
  unique()
gatunki_litX655 <- data.frame(gatunki = gatunki_litX655$X655, pole = "X655")

gatunki_litX650 <- data.frame(X650 = UKD_lit$X650) %>%
  unique() %>%
  filter(!is.na(X650))
gatunki_litX650$X650 <-str_replace_all(gatunki_litX650$X650,"(^|\\|)","~\\1")
gatunki_litX650 <- cSplit(gatunki_litX650,"X650",sep = "~",direction = "long")
gatunki_litX650 <- gatunki_litX650%>%
  filter(X650!="")
gatunki_litX650$X650 <- str_replace_all(gatunki_litX650$X650,"(.*?)(\\$a.*)(\\$2.*)","\\2")
gatunki_litX650 <- gatunki_litX650 %>%
  unique()
gatunki_litX650 <- data.frame(gatunki = gatunki_litX650$X650, pole = "X650")

gatunki_lit <- rbind(gatunki_litX655,gatunki_litX650)

gatunki_lit$gatunki <- str_remove_all(gatunki_lit$gatunki, ".*?\\$a")
gatunki_lit <- gatunki_lit[order(gatunki_lit$pole,gatunki_lit$gatunki),]
gatunki_lit <- gatunki_lit %>%
  unique()
#write.csv2(gatunki_lit, "C:/Users/User/Desktop/gatunki_literackie.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
#Basia definiuje, kt?re gatunki s? literackie i kt?rych gatunk?w nale?y szuka?

gatunki_lit <- gs_read(gs_title("gatunki_literackie_2005-2008"), ws = "gatunki_literackie")
gatunki_lit <- gatunki_lit %>%
  filter(`czy_gat_literacki?`=="tak")
gatunki_lit <- gatunki_lit %>%
  select(1,2) %>%
  unique()
bez_UKD_lit <- sqldf("select *
               from bn_ks_2005_2008 a
               where a.id not in 
               (select b.id
               from UKD_lit b)")
gatunki_lit_1 <- gatunki_lit %>%
  filter(pole=="X655") %>%
  select(1)
gatunki_lit_2 <- gatunki_lit %>%
  filter(pole=="X650") %>%
  select(1)

bez_UKD_lit_a <- sqldf("select *
                          from bez_UKD_lit a
                          join gatunki_lit_1 b on a.X655 like ('%'||b.gatunki||'%')")
bez_UKD_lit_b <- sqldf("select *
                          from bez_UKD_lit a
                          join gatunki_lit_2 b on a.X650 like ('%'||b.gatunki||'%')")

bez_UKD_lit <- sqldf("select *
                      from bez_UKD_lit a
                      where a.id in (select a.id from bez_UKD_lit_a a)
                      or a.id in (select a.id from bez_UKD_lit_b a)")
bez_UKD_lit <-bez_UKD_lit[order(bez_UKD_lit$X655,bez_UKD_lit$X600),]

#UKD_tfrtv
UKD_tfrtv <- mutate(bn_ks_2005_2008,
                    ukd_tfrtv = grepl("\\$a7\\.01", bn_ks_2005_2008$X080)|grepl("\\$a7\\.03", bn_ks_2005_2008$X080)|grepl("\\$a7\\.03\\(438\\)\\(091\\)", bn_ks_2005_2008$X080)|grepl("\\$a7\\.07", bn_ks_2005_2008$X080)|grepl("\\$a7\\.096", bn_ks_2005_2008$X080)|grepl("\\$a7\\.097", bn_ks_2005_2008$X080)|grepl("\\$a79", bn_ks_2005_2008$X080)|grepl("\\$a791", bn_ks_2005_2008$X080)|grepl("\\$a791\\.4", bn_ks_2005_2008$X080)|grepl("\\$a792", bn_ks_2005_2008$X080)|grepl("\\$a792\\.071", bn_ks_2005_2008$X080)|grepl("\\$a792\\.09", bn_ks_2005_2008$X080))
UKD_tfrtv <- UKD_tfrtv %>%
  filter(ukd_tfrtv == TRUE) %>%
  select(-length(UKD_tfrtv))
UKD_tfrtv <- sqldf("select *
               from UKD_tfrtv a
               where a.id not in 
               (select b.id
               from bez_UKD_lit b)
               and a.id not in 
               (select b.id
               from UKD_lit b)")

#bez_UKD_tfrtv
gatunki_tfrtvX655 <- data.frame(X655 = UKD_tfrtv$X655) %>%
  unique() %>%
  filter(!is.na(X655))

gatunki_tfrtvX655$X655<-str_replace_all(gatunki_tfrtvX655$X655,"(^|\\|)","~\\1")
gatunki_tfrtvX655<- cSplit(gatunki_tfrtvX655,"X655",sep = "~",direction = "long")
gatunki_tfrtvX655<- gatunki_tfrtvX655%>%
  filter(X655!="")
gatunki_tfrtvX655$X655 <- str_replace_all(gatunki_tfrtvX655$X655,"(.*?)(\\$a.*)(\\$2.*)","\\2")
gatunki_tfrtvX655 <- gatunki_tfrtvX655 %>%
  unique()
gatunki_tfrtvX655 <- data.frame(gatunki = gatunki_tfrtvX655$X655, pole = "X655")

gatunki_tfrtvX650 <- data.frame(X650 = UKD_lit$X650) %>%
  unique() %>%
  filter(!is.na(X650))
gatunki_tfrtvX650$X650 <-str_replace_all(gatunki_tfrtvX650$X650,"(^|\\|)","~\\1")
gatunki_tfrtvX650 <- cSplit(gatunki_tfrtvX650,"X650",sep = "~",direction = "long")
gatunki_tfrtvX650 <- gatunki_tfrtvX650%>%
  filter(X650!="")
gatunki_tfrtvX650$X650 <- str_replace_all(gatunki_tfrtvX650$X650,"(.*?)(\\$a.*)(\\$2.*)","\\2")
gatunki_tfrtvX650 <- gatunki_tfrtvX650 %>%
  unique()
gatunki_tfrtvX650 <- data.frame(gatunki = gatunki_tfrtvX650$X650, pole = "X650")

gatunki_tfrtv <- rbind(gatunki_tfrtvX655,gatunki_tfrtvX650)

gatunki_tfrtv$gatunki <- str_remove_all(gatunki_tfrtv$gatunki, ".*?\\$a")
gatunki_tfrtv <- gatunki_tfrtv[order(gatunki_tfrtv$pole,gatunki_tfrtv$gatunki),]
gatunki_tfrtv <- gatunki_tfrtv %>%
  unique()

#write.csv2(gatunki_tfrtv, "C:/Users/User/Desktop/gatunki_tfrtv.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
#Basia definiuje, kt?re gatunki s? tfrtv i kt?rych gatunk?w nale?y szuka?

gatunki_tfrtv <- gs_read(gs_title("gatunki_tfrtv_2005-2008"), ws = "gatunki_tfrtv")
gatunki_tfrtv <- gatunki_tfrtv %>%
  filter(`czy_gat_tfrtv?`=="tak")
gatunki_tfrtv <- gatunki_tfrtv %>%
  select(1,2) %>%
  unique()
bez_UKD_tfrtv <- sqldf("select *
               from bn_ks_2005_2008 a
               where a.id not in 
               (select b.id
               from UKD_tfrtv b)
               and a.id not in 
               (select b.id
               from bez_UKD_lit b)
               and a.id not in 
               (select b.id
               from UKD_lit b)")

gatunki_tfrtv_1 <- gatunki_tfrtv %>%
  filter(pole=="X655") %>%
  select(1)
gatunki_tfrtv_2 <- gatunki_tfrtv %>%
  filter(pole=="X650") %>%
  select(1)

bez_UKD_tfrtv_a <- sqldf("select *
                          from bez_UKD_tfrtv a
                          join gatunki_tfrtv_1 b on a.X655 like ('%'||b.gatunki||'%')")
bez_UKD_tfrtv_b <- sqldf("select *
                          from bez_UKD_tfrtv a
                          join gatunki_tfrtv_2 b on a.X650 like ('%'||b.gatunki||'%')")

bez_UKD_tfrtv <- sqldf("select *
                      from bez_UKD_tfrtv a
                      where a.id in (select a.id from bez_UKD_tfrtv_a a)
                      or a.id in (select a.id from bez_UKD_tfrtv_b a)")
bez_UKD_tfrtv <-bez_UKD_tfrtv[order(bez_UKD_tfrtv$X655,bez_UKD_tfrtv$X600),]

#reszta
reszta <- sqldf("select *
               from bn_ks_2005_2008 a
               where a.id not in 
               (select b.id
               from UKD_lit b)
               and a.id not in 
               (select b.id
               from bez_UKD_lit b)
               and a.id not in 
               (select b.id
               from UKD_tfrtv b)
               and a.id not in 
               (select b.id
               from bez_UKD_tfrtv b)")
#po??czenie z baz? PBL
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jdk1.8.0_191\\jre')
library(RJDBC)
jdbcDriver =JDBC("oracle.jdbc.OracleDriver",classPath="C:\\Users\\User\\Documents\\PBL\\ojdbc6.jar")
PBL <- dbConnect(jdbcDriver, "jdbc:oracle:thin:@//pbl.ibl.poznan.pl:1521/xe", "IBL_SELECT", "CR333444")
install.packages("pingr")
library(pingr)

gwiazdkowicze <- dbGetQuery(PBL,
                            "select *
                            from PBL_TWORCY tw where tw_.dbReadTable(PBL,'PBL_DZIALY')




write.csv2(UKD_lit, "C:/Users/User/Desktop/2005-2008_UKD_lit.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
write.csv2(bez_UKD_lit, "C:/Users/User/Desktop/2005-2008_bez_UKD_lit.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
write.csv2(UKD_tfrtv, "C:/Users/User/Desktop/2005-2008_UKD_tfrtv.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
write.csv2(bez_UKD_tfrtv, "C:/Users/User/Desktop/2005-2008_bez_UKD_tfrtv.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')
write.csv2(reszta, "C:/Users/User/Desktop/2005-2008_reszta.csv", row.names = F, na = '', fileEncoding = 'Windows-1250')

```

```{r}
#deleni poli na podpole
#field 600 <- number of field
#remember to change the name of the column from X600 to e.g. X787
marc_field <- cz_database %>%
  select(id,SIF,X600)%>%
  filter(X600!="")
marc_field$X600<-str_replace_all(marc_field$X600,"(^|\\|)","~\\1")
marc_field<- cSplit(marc_field,"X600",sep = "~",direction = "long")
marc_field<- marc_field%>%
  filter(X600!="")
marc_field$X600<-str_remove_all(marc_field$X600,"^\\|")
marc_field <- mutate(marc_field,
               indicator = str_replace_all(marc_field$X600,"(^.*?)(\\$.*)","\\1"))
subfield_list<- str_extract_all(cz_database$X600,"\\${2}.")
subfield_list<- unique(unlist(subfield_list))
empty_table<- data.frame(matrix(ncol = length(subfield_list),nrow = lengths(marc_field)[1]))
colnames(empty_table) <-subfield_list

marc_field<-cbind(marc_field,empty_table)

subfield_list_char <- paste("(",subfield_list,")",sep = "")
subfield_list_char <- str_replace_all(subfield_list_char,"\\$","\\\\$")

x <- 1:length(subfield_list)

for (i in x) {
  marc_field$X600 <- str_replace(marc_field$X600,subfield_list_char[i],"|\\1")
  progress(match(i,x), max.value = length(x)) 
}
for (i in x) {
  subfield_list_char2 <- str_replace_all(subfield_list,"\\$","\\\\$")
string_a <- "(^)(.*?\\|"
string_b <- subfield_list_char2[i]
string_c <- ")(.*?)(\\,{0,1})((\\|\\${2})(.*)|$)"
string <- paste(string_a,string_b,string_c,sep = "")
marc_field[,i+4] <- ifelse(grepl(subfield_list_char2[i],marc_field$X600),str_replace_all(gsub(string,"\\3",marc_field$X600),"\\${2}.", "~"),NA)
progress(match(i,x), max.value = length(x))  
}
```
